---
title: "Data 607 Final Project"
author: "Joshua Registe"
date: "12/11/2019"
output:
  html_document:
    #code_folding: hide
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
---


# Motivation
<br><br>

As part of this final project, I will be looking at weather patterns globally over the past few decades looking at potential major contributors to the cause of global warming. Engineers and scientists often get posed questions along the lines of infrastructure resiliency and protection against the effects of storms, rising sea level, air quality and water resources, all which are increasingly seeing the effects of climate change. It is important to understand his and be able to propose solutions that are feasible for the future of the planet.<br><br>


```{r, echo=FALSE}

url <- "https://ane4bf-datap1.s3-eu-west-1.amazonaws.com/wmocms/s3fs-public/styles/featured_media_detail/public/news/featured_media/new-climate-statement-infographic_2018_0.png?lkfmoahuwvswsRsKZHRa2FMazya_Asrb&itok=azeHuK21"
```

<center><img src="`r url`"></center>
<br><br>

Some of the data sources that will be considered include:

data.world climate data, kaggle greenhouse gas dataset containing data on containing global population, emissions, temperature, energy, etc. As well as research article data from the Sciencenews website. 

data sources:<br>
<https://data.world/worldbank/climate-change-data><br>
<https://www.kaggle.com/unitednations/international-greenhouse-gas-emissions/data#><br>
<https://www.sciencenews.org/><br><br>


The presentation will follow OSEMN workflow for obtaining data, scrubbing data, exploring data, modeling data, and interpreting results

# Library Preparation

```{r setup,warning=FALSE,message=FALSE}
library(readxl)
library(dplyr)
library(tidyr)
library(stringr)
library(tidyverse)
library(httr)
library(rvest)
library(digest)
library(tidytext)
library(wordcloud)
library(RColorBrewer)
library(XML)
library(stringr)
library(ggplot2)
library(tm)
library(magrittr)
library(rworldmap)
library(caret)
library(rpart)
library(rpart.plot)
library(ggcorrplot)
library(Hmisc)
library(DT)
library(knitr)
```


# Qualitative Analysis
## Web Scraping
```{r, echo=FALSE}

url <- "https://miro.medium.com/max/1132/1*YfeP5WFbn0MwI76kuTM38w.png"
```

<center><img src="`r url`"></center>
<br><br>


To begin, Web scraping was done with the following script to pull data from <https://www.sciencenews.org/>. This specific script allowed for looping through all pages of science news under the search query of "Climate Change" and "Global Warming". 

```{r, echo=FALSE}
#setwd('C:/Users/joshu/OneDrive - CDM Smith/Documents/CUNY SPS/Data 607/Assignments/Final Project')
setwd('C:/Users/registejh/OneDrive - CDM Smith/Documents/CUNY SPS/Data 607/Assignments/Final Project')
```




```{r, eval=FALSE}
ArticleListings <- data.frame(title=character(),
                              summary=character(), 
                              link=character(), 
                              description = character(),
                              stringsAsFactors=FALSE) 

#causes of climate change searched on Science news
for (i in seq(0,40,1)){
  
  var<-read_html(paste0('https://www.sciencenews.org/page/',i,'?s=cause+of+climate+change&topic=&start-date=&end-date=&orderby=relevance'))
  title<-html_nodes(var,'.post-item-river__title___J3spU a') %>% html_text(trim = TRUE)
  link<-html_nodes(var,'.post-item-river__title___J3spU a') %>% html_attr('href')
  summary<-html_nodes(var,'.post-item-river__excerpt___3ok6B') %>% html_text(trim = TRUE)
  
  
  ArticleListingstemp <- data.frame(title=character(),
                                    summary=character(), 
                                    link=character(),
                                    stringsAsFactors=FALSE) 
  
  ArticleListingstemp <- rbind(ArticleListingstemp, as.data.frame(cbind(title,
                                                                        summary,
                                                                        link)))
  
  ArticleListings<- bind_rows(ArticleListings,ArticleListingstemp)
  
}

#causes of global warming searched on science articles

for (i in seq(0,40,1)){
  
  var<-read_html(paste0('https://www.sciencenews.org/page/',i,'?s=cause+of+global+warming&topic=&start-date=&end-date=&orderby=relevance'))
  title<-html_nodes(var,'.post-item-river__title___J3spU a') %>% html_text(trim = TRUE)
  link<-html_nodes(var,'.post-item-river__title___J3spU a') %>% html_attr('href')
  summary<-html_nodes(var,'.post-item-river__excerpt___3ok6B') %>% html_text(trim = TRUE)
  
  ArticleListingstemp <- data.frame(title=character(),
                                    summary=character(), 
                                    link=character(),
                                    stringsAsFactors=FALSE) 
  
  ArticleListingstemp <- rbind(ArticleListingstemp, as.data.frame(cbind(title,
                                                                        summary,
                                                                        link)))
  
  ArticleListings<- bind_rows(ArticleListings,ArticleListingstemp)
  }

for (i in 1:nrow(ArticleListings)){
  
  ArticleListings[i,4]<- read_html(ArticleListings[i,3])%>%html_nodes('.single__rich-text___BlzVF p')%>% html_text()%>%paste(collapse = " ")
}
rm(ArticleListingstemp)
```



```{r, echo=FALSE}
ArticleListings<-read.csv('articlelistings.csv')

```




```{r}

datatable(ArticleListings[c(1:4)])

```


We can create a function as shown in the following code to allow us to do a word cloud on any of the fields from dataset's column. With our webscraped dataset from <https://www.sciencenews.org/>, 3 word clouds were generated for the title of the articles, the summary of the articles, and the articles themselves from every link extracted into the dataframe. This allows us to visualize the patterns in words associated with "climate change" and "global warming" searches. Within these word clouds, sentimental words (words that describe feelings and emotions), and stop words (words that appear frequently such as "a" in the english language). 

## Text Visualization

```{r,message=FALSE,warning=FALSE}
wordextract<- function(dfcolumn){
descriptionwords <- dfcolumn %>% str_replace_all("^<[:graph:]*>$","")%>%
  str_replace_all("\\\n"," ") %>%
  str_replace_all("[^[:alpha:] ]"," ") %>% str_replace_all("\\s[:alpha:]{1,2}\\s","")%>% tolower()

#pulling all individual words into a dataframe and counting
word_counts <- as.data.frame(table(unlist(strsplit(descriptionwords, "\\s+"))))
colnames(word_counts)<-c("word","Freq")
#removing filler words and sentimental words
word_counts <-anti_join(word_counts,get_stopwords())%>%
  anti_join(get_sentiments())
wordcloud(words = word_counts$word, freq = word_counts$Freq, min.freq = 1,
          max.words=100, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))

}
```



```{r,message=FALSE,warning=FALSE}
wordextract(ArticleListings$title)
wordextract(ArticleListings$summary)
wordextract(ArticleListings$description)
```

<br>
We can clearly see some of the words associated with climate change such as: emissions, sea level rise, carbon, earth, ocean, environment, arctic, oceans, droughts, tides, etc. these words represent both causes and effects of climate change.

<br>
<br>

# Quantitative Analyses 1 

## Data Collection

```{r, echo=FALSE}

url <- "https://www.smartdatacollective.com/wp-content/uploads/2019/09/data-collection-ideas-1024x711.jpg"
```

<center><img src="`r url`"></center>
<br><br>


Diving deeper into more quantitative analyses, a dataset from data.world was extracted look at the greenhouse gas emissions world wide. 


<https://data.world/worldbank/climate-change-data>

```{r, message=FALSE, warning=FALSE}

GHG_Inventory<- read_csv("greenhouse_gas_inventory_data_data.csv")
GHG_unique_cat<-unique(GHG_Inventory$category)
GHG_unique_loc<-unique(GHG_Inventory$country_or_area)
GHG_unique_cat

```


## Data Cleansing

```{r, echo=FALSE}

url <- "https://miro.medium.com/max/500/1*yWFQiGjlgHUVYeh4ELELyw.jpeg"
```

<center><img src="`r url`"></center>
<br><br>

Using the Dplyr package and the stringr package, we can deploy various techniques to treating this data into a more usable and readable format. The following script will adjust the unique names in each column so that they are shortened and better represented graphically. Additionally, we can mutate this information and group by the countries within our dataset to see where the percentage of total emissions lie by country as shown in the following code.

```{r}
{
GHG_Inventory$category<- GHG_Inventory$category %>% str_replace_all(GHG_unique_cat[[2]],"GHG_Emissions")
GHG_Inventory$category<- GHG_Inventory$category %>% str_replace_all(GHG_unique_cat[[1]],"CO2_Emissions")
GHG_Inventory$category<- GHG_Inventory$category %>% str_replace_all(GHG_unique_cat[[3]],"GHGEmissions_W/O_LULUCF")
GHG_Inventory$category<- GHG_Inventory$category %>% str_replace_all(GHG_unique_cat[[4]],"HFC_Emissions")
GHG_Inventory$category<- GHG_Inventory$category %>% str_replace_all(GHG_unique_cat[[5]],"CH4_Emissions")
GHG_Inventory$category<- GHG_Inventory$category %>% str_replace_all(GHG_unique_cat[[6]],"NF3_Emissions")
GHG_Inventory$category<- GHG_Inventory$category %>% str_replace_all(GHG_unique_cat[[7]],"N2O_Emissions")
GHG_Inventory$category<- GHG_Inventory$category %>% str_replace_all(GHG_unique_cat[[8]],"PFC_Emissions")
GHG_Inventory$category<- GHG_Inventory$category %>% str_replace_all(GHG_unique_cat[[9]],"SF6_Emissions")
GHG_Inventory$category<- GHG_Inventory$category %>% str_replace_all(GHG_unique_cat[[10]],"HFC\\+PFC")
}
GHG_unique_cat<-unique(GHG_Inventory$category)
GHG_unique_cat
GHG_Inventoryfreq<-GHG_Inventory %>%filter(category==GHG_unique_cat[[1]]) %>% 
  group_by(country_or_area)%>%
  mutate(percbyloc=value/sum(value))
```


Because this is geographic data, this is best visualized spatially and we can use the rworld package within r to map our dataset. minor tweaks to our original dataset were made so that 1 to 1 matches were obtained for all countries where data exists. A visualisation is shown below of the global contributors of climate change. 


```{r}

map.world<-map_data("world")

GHG_Inventoryfreq$country_or_area<- str_replace_all(GHG_Inventoryfreq$country_or_area,"Russian Federation","Russia" )
GHG_Inventoryfreq$country_or_area<- str_replace_all(GHG_Inventoryfreq$country_or_area,"United Kingdom","UK" )
GHG_Inventoryfreq$country_or_area<- str_replace_all(GHG_Inventoryfreq$country_or_area,"United States of America","USA" )
map.world2<-left_join(map.world, GHG_Inventoryfreq, by=c('region'='country_or_area'))

ggplot()+
 theme(legend.title = element_text("Percent"))+
 geom_map(data=map.world2, map=map.world2, aes(map_id=region, x=long, y=lat, fill=percbyloc))+
 scale_fill_gradient(low = "lightblue",high = "purple", guide = "colourbar")+
 coord_equal()+ 
  labs(x = "Latitude",
       y = 'Longitude',
       title = 'Percent of Global CO2 Emissions by Country')
```

<br>
Divinga little deeper into this dataset, the dataframe was filtered for just the United States and a visualisation of the emissions over time are shown in the following figure. It is very clear how within the past decade, many of the greenhouse gases (and more notably, the anthropogenic ones such as CO2) have been increasing, and the rate at which its been increasing with it. CH4 which has high GHG potential is shown to be decreasing overtime.
<br>


```{r}

GHG_Inventory%>% 
  filter(country_or_area==GHG_unique_loc[1])%>%
  ggplot()+
  geom_point(mapping = aes(year,value, color = category))+geom_smooth(mapping = aes(year,value, color = category))+
  facet_wrap(category~.,scales='free_y')+
  
  theme(strip.text.x=element_blank(),legend.position = 'bottom', axis.text.x = element_text(angle = 90))+
  labs(title = "U.S. Greenhouse Gas Emissions Over Time",
       x = "Eq Metric Tons CO2",
       y = "Year")
```

# Quantitative Analyses 2

## Data Collection and Cleaning

<br>
To dive deeper into the causes of climate change, a dataset from 
<https://www.kaggle.com/unitednations/international-greenhouse-gas-emissions/data#> was used which has has information on population, greenhouse gas emissions, energy, temperature and more. 
<br>

This dataset in its native form was extracted in a wide format with years as columns. To perform the analyses we intend, we will need to get our data into a long format for the years, and then re-spread the data based on "Series Name" which represent the potential causes of climate change. This spreading of the data will allow us to compute pairwise comparisons greenhouse gas emissions to other factors.

<br>

```{r, warning=FALSE}

climatechange<-read_excel("climate_change_download_0.xls")

#convert to long
climatechange_gather<-climatechange%>%gather(key = "DATE",value = "VALUE",-(1:6),(7:ncol(climatechange)))

climatechangeunique<-unique(climatechange_gather$`Series name`)
climatechange_gather<-climatechange_gather[c(2,4,7,8)]
climatechange_gather$VALUE<-as.numeric(climatechange_gather$VALUE)

ClimatechangeUS<-climatechange_gather%>%filter(`Country name`=="United States")
```

<br>
The first analysis that was done was a dendrogram computing a classification of CO2 emissions based on the other 60 parameters in the dataset. CO2 emissions was converted into a classification problem where quartiles were used to provide level of risk associated with climate change:
<br>1 - Low risk
<br>2 - Low/medium risk
<br>3 - Medium/high risk
<br>4 - High risk

<br>
Using the carat package, we increased the number of samples within our dataset by randomly sampling with replacement from our original dataset to have 1000 observations. additionally, we removed parameters that do not have enough variance to provide meaningful results were removed from the dataset. 
<br>
From the dendrogram, we can see that R has selected approximately 5/60 of our paramater as our major contributors of climate change, and of those 5, energy use per capita was had the highest risk for impacting global warming.
<br>

```{r}

ClimatechangeUSTree<- ClimatechangeUS %>% spread(`Series name`,VALUE)


ClimatechangeUSTree$CO2Thresh<-as.factor(
ntile(ClimatechangeUSTree$`CO2 emissions, total (KtCO2)`, 4))


ClimatechangeUSMatrix<-ClimatechangeUSTree[c(-1,-2,-13,-14)]

ClimatechangeUSTree<-ClimatechangeUSTree[c(-1,-2,-13,-14,-15)]


ClimatechangeUSTree <- ClimatechangeUSTree[,colSums(is.na(ClimatechangeUSTree))<nrow(ClimatechangeUSTree)-5]

ClimatechangeUSTree<- ClimatechangeUSTree%>%filter(is.na(CO2Thresh)==FALSE)

ClimatechangeUSTreeS<- sample_n(ClimatechangeUSTree,1000,replace= TRUE)

SingleTree <- rpart(formula = CO2Thresh ~ ., 
                    data = ClimatechangeUSTreeS,
                    method = 'class')

rpart.plot(x = SingleTree, yesno = 2, type = 5,extra = 108,tweak = 1, shadow.col = "lightgrey")
```

<br>
Finally,  We can write a function to create a pairwise correlation matrix of all the parameters in our dataset and look at spearman regression coefficients to see which of our parameters have strong monotonic (increasing or decreasing coincidentally) relationships with CO2 emissions.  We can see from the figure below that some of the parameters with the strongest relationship to CO2 emissions include:
<br>Energy consumption
<br>Population
<br>

```{r}


Pearson_Matrix<-function(Data,siglevel,statmethod,labsize){
  #get all the numeric data out for meatrix
  Data<-select(Data,which(sapply(Data,class)=="numeric"))
  cor_data<- rcorr(as.matrix(Data),type = statmethod)
  M <- cor_data$r
  #p to designate significant correlations
  p_mat <- cor_data$P
  
  for(i in 1:ncol(Data)){
    try(p_mat[i,i]<-0)
  }
    Matrix<-ggcorrplot(M,method = "square", colors = c("#E46726", "white","#6D9EC1"),lab = TRUE, 
                     lab_size = labsize,type = "full", p.mat=p_mat, sig.level = siglevel,legend.title = statmethod,
                     title = paste0(statmethod," Correlation Matrix" ))
  print(Matrix)
}
```



```{r,fig.height=16, fig.width=18}
ClimatechangeUSMatrix<- ClimatechangeUSMatrix[-length(ClimatechangeUSMatrix)]
ClimatechangeUSMatrix <- ClimatechangeUSMatrix[,colSums(is.na(ClimatechangeUSMatrix))<nrow(ClimatechangeUSMatrix)-15]

ClimatechangeUSTreeS<- sample_n(ClimatechangeUSTree,100,replace= TRUE)
Pearson_Matrix(ClimatechangeUSMatrix,.05,"spearman",5)
```


# Conclusion

There is an extensive library on all things related to climate change and this project barely scratches the surface, however we are able to portray just how valuable some of the tools within R can be for:
<br>
Webscraping - Using packages such as rvest to perform text-based analyses on science related articles on the subject of interest.
<br>
Data collection and cleaning - Using packages such as dplyr, tidyr and stringr to to maniputlate data in a form that would allow us perform certain analyses and visualisations
<br>
Analysis and Visualisation - Using packages within the tidyverse to perform highly customizable visualisations as well as packages Hmisc and r.part and ggcorrplot to perform different pairwise analyses.
<br>
We were able to look into many of the contributing factors associated with greenhouse gas emissions and climate change as well as come up with preliminary assessments on level of impacts for all of these parameters.
