---
title: "Week 7 Assignment"
author: "Joshua Registe"
date: "10/13/2019"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries}
library(rvest)
library(xml2)
library(jsonlite)
library(XML)
library(kableExtra)
```


Files for different datatypes from chapters 2-4 of "Automated Data Collection" have been created and can be found on github here in json, xml, and html formats:
<br>
XML:
<br>
<https://raw.githubusercontent.com/joshuargst/Data607Week7/master/Books.xml>
<br>
HTML:
<br>
<https://raw.githubusercontent.com/joshuargst/Data607Week7/master/Books.html>
<br>
json:
<br>
<https://raw.githubusercontent.com/joshuargst/Data607Week7/master/Books.json>
<br>

# XML

The XML file is concocted as shown below:
```{r }

books.xml<-read_xml("https://raw.githubusercontent.com/joshuargst/Data607Week7/master/Books.xml") %>%xmlParse()
books.xml

```



The tables can be extracted via the "XML" package in r and the dataframe can be extracted as such:
```{r}

kable(xmlToDataFrame(books.xml))
```


# HTML

The HTML file is concocted as shown below:
```{r}
Books.html<- read_html("https://raw.githubusercontent.com/joshuargst/Data607Week7/master/Books.html") 
Books.html
```

tables from the file can be extracted via the "rvest" package in r as shown below:
```{r}
Books.html%>% html_table()%>% kable()
```

# JSON

Tables from json files can be extracted via the "jsonlite" package and is shown in the following format below:
```{r}
read_json("https://raw.githubusercontent.com/joshuargst/Data607Week7/master/Books.json")

```


From within the jsonlite package, there are also ways to extract tabular information to be parsed out as shown below:
```{r}
kable(fromJSON("https://raw.githubusercontent.com/joshuargst/Data607Week7/master/Books.json"))
```
<br>
<br>
<br>
<br>
When using the proper packages and information parsers, the 3 dataframes come out identical despite what format the data originated from.

