---
title: "Week 9 Assignment"
author: "Joshua Registe"
date: "10/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 9 Assignment

The following libraries are required to extract the information from the NYtimes API and convert this to a dataframe

```{r, message=FALSE, warning=FALSE}
library(httr)
library(jsonlite)
library(knitr)
library(DT)

```


We can extract the URL from the NYtimes website. An account was created to apply and obtain an API key from NYTimes devs and allow for the extraction of all data.

```{r}

URL <- "https://api.nytimes.com/svc/movies/v2/reviews/all.json"

Movie_reviewsAPI <- GET(URL, query = list("api-key" = "seSiAqKUdWfXFqJTiQRQBTuBMRgiTYU6"))
APIContent <- content(Movie_reviewsAPI, "text")

```

We can use the jsonlite package to extract information from our url and convert the json information into a dataframe as shown below.

```{r}
Movie.Information <- fromJSON(APIContent)
Movie.Information<- as.data.frame(Movie.Information$results)

datatable(Movie.Information)

```

