---
title: "606_Presentation"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

An airline charges the following baggage fees: $25 for the first bag and $35 for the
second. Suppose 54% of passengers have no checked luggage, 34% have one piece of checked luggage and
12% have two pieces. We suppose a negligible portion of people check more than two bags.

(a) Build a probability model, compute the average revenue per passenger, and compute the corresponding
standard deviation.



```{r cars}
Airlineevent<-data.frame("Bag_Count"= c(0,1,2),
                        "Baggage_Fee" = c(0,25,35), 
                        "Percent_Passengers" = c(0.54,0.34,0.12))

Airlineevent
```




```{r Revenue_Per_Passenger}
Revenue_per_Pas <- sum(Airlineevent$Baggage_Fee*Airlineevent$Percent_Passengers)

print(Revenue_per_Pas)


```






```{r }

Revenue_Variance<-sum(Airlineevent$Percent_Passengers*(Airlineevent$Baggage_Fee- Revenue_per_Pas)^2)

Revenue_Stdev<-sqrt(Revenue_Variance)

print(Revenue_Stdev)

```



(b) About how much revenue should the airline expect for a 
ight of 120 passengers? With what standard deviation? Note any assumptions you make and if you think they are justifed.



```{r}
num_Pas<-120
estimated_revenue <- num_Pas*Revenue_per_Pas
print(estimated_revenue)

estimated_devation<- sqrt(Revenue_Variance*num_Pas)
print(estimated_devation)

```




